# Missing values

```{r}
library(tidyverse)
taxi.09<-read_csv("https://s3.amazonaws.com/nyc-tlc/trip+data/yellow_tripdata_2009-01.csv")
taxi.12<-read_csv("https://s3.amazonaws.com/nyc-tlc/trip+data/yellow_tripdata_2012-01.csv")

set.seed(0)
sample09 <-sample_n(taxi.09,1000000, replace = FALSE)
sample12 <-sample_n(taxi.12,1000000, replace = FALSE)

```

```{r}

taxi5<-read_csv("https://s3.amazonaws.com/nyc-tlc/trip+data/yellow_tripdata_2014-05.csv")
set.seed(0)
sample_taxi5<-sample_n(taxi5,652435, replace = FALSE)

missing.values <- sample_taxi5 %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <-
    (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() +
        geom_bar(aes(x = reorder(key, desc(pct)), 
                     y = pct, fill=isna), 
                 stat = 'identity', alpha=0.8) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", 
                        values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() +
      labs(title = "Percentage of missing values", x =
             'Variable', y = "% of missing values")

percentage.plot 

```
In the sampling dataset of Yellow Taxi Trip May 2014, as we can see from this "Percentage of missing values" plot, the only variable contains NAs are "store_and_fwd_flag" and around 50% of the values are missing.


```{r}

missing.values <- sample09 %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <-
    (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() +
        geom_bar(aes(x = reorder(key, desc(pct)), 
                     y = pct, fill=isna), 
                 stat = 'identity', alpha=0.8) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", 
                        values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() +
      labs(title = "Percentage of missing values", x =
             'Variable', y = "% of missing values")

percentage.plot 

```


In the sampling dataset of Yellow Taxi Trip January 2009, as we can see from this "Percentage of missing values" plot, all values of these three variables are missing.


```{r}

missing.values <- sample12 %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <-
    (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() +
        geom_bar(aes(x = reorder(key, desc(pct)), 
                     y = pct, fill=isna), 
                 stat = 'identity', alpha=0.8) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", 
                        values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() +
      labs(title = "Percentage of missing values", x =
             'Variable', y = "% of missing values")

percentage.plot 

table(colSums(is.na(sample12)))



```
In the sampling dataset of Yellow Taxi Trip January 2012, as we can see from this "Percentage of missing values" plot, almost 50% values of "store_and_fwd_flag" are missing and only very few values of "dropoff_latitude" and "dropoff_longitude" are missing.  